Typed Clojure

### 2014-10-04 icfp15-typedclojure.tex
add 7 place relation and some color

\usepackage{mmm}
\usepackage{mathpartir}
\usepackage{clj-grammar}
%\usepackage{url}
%\usepackage{biblatex}
\include{bibliography.bib}

@@

We describe our type system in a similar way to Typed Racket (\citet{TF10}).
The main judgement 
{\judgement {\Gamma} {\hastype {\e{}} {\t{}}} 
  {\filterset {\thenprop {\prop{}}}
              {\elseprop {\prop{}}}}
            {\object{}}}
says expression \e{} is of type \t{} in environment $\Gamma$, with 
then proposition {\thenprop {\prop{}}}, else proposition {\elseprop {\prop{}}}
and object \object{}.

@@

\begin{figure*}
$$
\begin{altgrammar}
  \expd{}, \e{} &::=& \x\ \alt {\comb {\e{}} {\e{}}} \alt {\abs {\x a} {\e{}}}
                      \alt \true{} \alt \false{} \alt \nil{}
                      \alt \class{}
                      \alt {\tryexp {\e{}} {\class} {\x} {\e{}} {\e{}}}
                      \alt {\ifexp {\e{}} {\e{}} {\e{}}}
                      \\ &\alt& 
                      {\doexp {\e{}} {\e{}}}
                      \alt
                      {\fnexp {\overrightarrow{\hastype {\x{}} {\t{}}}} {\e{}}}
                      \alt {\letexp {\x{}} {\e{}} {\e{}}}
                      \alt {\fieldexp {fld} {\e{}}}
                      \alt {\methodexp {fld} {\e{}} {\overrightarrow{\e{}}}}
                      \alt {\newexp {\e{}} {\overrightarrow{\e{}}}}
                &\mbox{Expressions} \\ \\
  \s{}, \t{}    &::=& \Top \alt (\class{}\ \overrightarrow{\t{}})
                      \alt \class\ 
                      \alt \protocol\ 
                      \alt {\Instance {\class}}
                      \alt {\Satisfies {\protocol}}
                      \alt {\Value \v{}}
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt {\In {\t{}\ \overrightarrow{\t{}}}}
                      \\ &\alt&
                      %\alt (\Not\ \t{})
                      {\All {\overrightarrow{\x{}}} {\t{}}}
                      \alt
                      {\TFn {\overrightarrow{\x{}}} \t{}}
                      \alt
                      {\Arrow {\overrightarrow{ \hastype {\x{}} {\t{}}}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                      \alt
                      {\ArrayTwo {\t{}} {\t{}}}
                      \alt {\HMapc {\mandatory}}
                      \alt {\HMapp {\mandatory} {\absent}}

@@

\begin{figure*}
\begin{mathpar}
\infer [T-Local]
{ \inpropenv {\propenv{}} {\isprop {\x{}} {\t{}}}}
{ \judgement {\propenv{}} {\hastype {\x{}} {\t{}}} {\filterset {{\notprop {\falsy} {\x{}}}} {{\isprop {\falsy} {\x{}}}}}
             {\x{}}}

\infer [T-True]
{}
{ \judgement {\propenv{}} {{\hastype {\true{}} {\True{}}}} {{\filterset {\topprop{}} {\botprop{}}}} {\emptyobject{}} }

\infer [T-False]
{}
{ \judgement {\propenv{}} {{\hastype {\false{}} {\False{}}}} {\filterset {\botprop{}} {\topprop{}}} {\emptyobject{}} }

\infer [T-Nil]
{}
{ \judgement {\propenv{}} {{\hastype {\nil{}} {\Nil{}}}} {\filterset {\botprop{}} {\topprop{}}} {\emptyobject{}} }

\infer [T-Do]
{ \judgement {\propenv{}} {\hastype {\e1} {\t1}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop1}}} {\object{1}} 
\\\\
  \judgement {\propenv{}, {\orprop {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}}
                 {\hastype {\e{}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}} 
                 {\object{}}
  }
{ \judgement {\propenv{}} {\hastype {\doexp {\e1} {\e{}}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}} {\object} 
             }

\infer [T-If]
{ \judgement {\propenv{}} {\hastype {\e1} {\t{1}}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}
                 {\object{1}} 
             \\\\
  \judgement {\propenv{}, {\thenprop {\prop{1}}}}
                 {\hastype {\e2} {\t{}}} {\filterset {\thenprop {\prop{2}}} {\elseprop {\prop{2}}}}
                 {\object{}}
  \\\\
  \judgement {\propenv{}, {\elseprop {\prop{1}}}}
                 {\hastype {\e3} {\t{}}} {\filterset {\thenprop {\prop{3}}} {\elseprop {\prop{3}}}}
                 {\object{}}
             }
{ \judgement {\propenv{}} {\hastype {\ifexp {\e1} {\e2} {\e3}} {\t{}}} 
                 {\filterset {\orprop {\thenprop {\prop{2}}} {\thenprop {\prop{3}}}} 
                             {\orprop {\elseprop {\prop{2}}} {\elseprop {\prop{3}}}}}
                 {\object{}} }

\infer [T-Let]
{ \judgement {\propenv{}} {\hastype {\e{1}} {\s{}}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}
             {\object{1}}
  \\\\
  \judgement {\propenv{}, {\isprop {\s{}} {\x{}}}, 
              {\impprop {\notprop {\falsy} {\x{}}} {\thenprop {\prop{1}}}}, 
              {\impprop {\isprop {\falsy} {\x{}}} {\elseprop {\prop{1}}}}}
             {\hastype {\e{}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
             {\object{}} 
             }
{ \judgement {\propenv{}} {\hastype {\letexp {\x{}} {\e{1}} {\e{}}} {\t{}}}
             {\replacefor {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                          {\object{1}}
                          {\x{}}}
             {\replacefor {\object{}} 
                          {\object{1}}
                          {\x{}}}
             }

\infer [T-Throw]
{ \judgement {\propenv{}} {\hastype {\e{}} {\Throwable}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
}
{ \judgement {\propenv{}} {\hastype {\throwexp {\e{}}} {\Bottom}} {\filterset {\botprop{}} {\botprop{}}}
             {\emptyobject{}} }

\infer [T-App]
{ \judgement {\propenv{}} {\hastype {\e{}} {\Arrow {\hastype {\x{}} {\s{}}}
                                                       {\t{}}
                                                       {\filterset {\thenprop {\prop{f}}}
                                                                   {\elseprop {\prop{f}}}}
                                                       {\object{f}}}}
                {\filterset {\thenprop {\prop{}}}
                            {\elseprop {\prop{}}}}
                {\object{}}
  \\\\
  \judgement {\propenv{}}
                 {\hastype {\ep{}} {\s{}}}
                 {\filterset {\thenprop {\propp{}}}
                             {\elseprop {\propp{}}}}
                 {\objectp{}} 
}
{ \judgement {\propenv{}} {\hastype {\appexp {\e{}} {\ep{}}}
                                        {\replacefor {\t{}}
                                                     {\objectp{}}
                                                     {\x{}}}}
                 {\replacefor {\filterset {\thenprop {\prop{f}}}
                                          {\elseprop {\prop{f}}}}
                              {\objectp{}}
                              {\x{}}}
                 {\replacefor {\object{f}}
                              {\objectp{}}
                              {\x{}}}
}

\infer [T-Abs]
{ \judgement {\propenv{}, {\isprop {\s{}} {\x{}}}}
             {\filterset {\thenprop {\prop{}}}
                         {\elseprop {\prop{}}}}
             {\object{}}}
{ \judgement {\propenv{}} {\hastype {\e{}} {\ArrowOne {\x{}} {\s{}}
                                                      {\t{}}
                                                      {\filterset {\thenprop {\prop{}}}
                                                                  {\elseprop {\prop{}}}}
                                                      {\object{}}}}
                {\filterset {\topprop}
                            {\botprop}}
                {\object{}}}

\infer [T-Try]
{ \judgement {\propenv{}} {\hastype {\e{1}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
\\\\
  \judgement {\propenv{}, {\isprop {\x{}} {\class{1}}}}
             {\hastype {\e{2}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
\\
  \judgement {\propenv{}}
             {\hastype {\e{3}} {\t{3}}}
                 {\filterset {\thenprop {\prop{3}}} {\elseprop {\prop{3}}}}
                 {\object{3}}
}
{ \judgement {\propenv{}} {\hastype {\tryexp {\e{1}} {\class{1}} {\x{}} {\e{2}} {\e{3}}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
                 }

\infer [T-New]
{ \ctorparams {\overrightarrow {\e{i}}} {\overrightarrow {\t{i}}}
  \\
  {\overrightarrow{
  \judgement {\propenv{}} {\hastype {\e{i}} {\t{i}}} 
             {\filterset {\prop{}} {\prop{}}}
         {\object{}}}}^{i}
           }
{ \judgement {\propenv{}} {\hastype {\newexp {\class{}} {\overrightarrow{\e{i}}}} {\class{}}}
             {\filterset {\topprop{}} {\botprop{}}}
             {\emptyobject}
             }

@@

\infer [T-Subsume]
{ \judgement {\propenv{}} {\hastype {\e{}} {\t{}}}
             {\filterset {\thenprop {\prop{}}}
                         {\elseprop {\prop{}}}}
             {\object}
\\\\
\inpropenv {\propenv{}, {\thenprop {\prop{}}}} {\thenprop {\propp{}}}
\\
\inpropenv {\propenv{}, {\elseprop {\prop{}}}} {\elseprop {\propp{}}}
\\\\
\inpropenv {} {\issubtype {\t{}} {\tp{}}}
\\
\inpropenv {} {\issubtype {\object{}} {\objectp{}}}
}
{ \judgement {\propenv{}} {\hastype {\e{}} {\tp{}}}
             {\filterset {\thenprop {\propp{}}}
                         {\elseprop {\propp{}}}}
             {\objectp}}

>>>>

\usepackage{mmm}
\usepackage{mathpartir}
\usepackage{clj-grammar}
\usepackage{xcolor}
%\usepackage{url}
%\usepackage{biblatex}
\include{bibliography.bib}

@@

We describe our type system in a similar way to Typed Racket (\citet{TF10}).
The main judgement 
{\judgementsevcol {\mmenv}
                 {\taenv}
                 {\propenv}
               {\hastype {\e{}} {\t{}}}
  {\filterset {\thenprop {\prop{}}}
              {\elseprop {\prop{}}}}
  {\object{}}
  {\mmenvp}}
says expression \e{} is of type \t{} in the 
multimethod environment $\mmenv$, the \emph{tools.analyzer} environment
$\taenv'$
and proposition environment $\propenv$, with 
then proposition {\thenprop {\prop{}}}, else proposition {\elseprop {\prop{}}},
object \object{} and new multimethod environment
${\mmenvp}$.

@@

\begin{figure*}
$$
\begin{altgrammar}
  \expd{}, \e{} &::=& \x\ \alt {\comb {\e{}} {\e{}}} \alt {\abs {\x a} {\e{}}}
                      \alt \true{} \alt \false{} \alt \nil{}
                      \alt \class{}
                      \alt {\tryexp {\e{}} {\class} {\x} {\e{}} {\e{}}}
                      \alt {\ifexp {\e{}} {\e{}} {\e{}}}
                      \\ &\alt& 
                      {\doexp {\e{}} {\e{}}}
                      \alt
                      {\fnexp {\overrightarrow{\hastype {\x{}} {\t{}}}} {\e{}}}
                      \alt {\letexp {\x{}} {\e{}} {\e{}}}
                      \alt {\fieldexp {fld} {\e{}}}
                      \alt {\methodexp {mth} {\e{}} {\overrightarrow{\e{}}}}
                      \alt {\newexp {\class{}} {\overrightarrow{\e{}}}}
                &\mbox{Expressions} \\ \\
  \s{}, \t{}    &::=& \Top \alt (\class{}\ \overrightarrow{\t{}})
                      \alt \class\ 
                      \alt \protocol\ 
                      \alt {\Instance {\class}}
                      \alt {\Satisfies {\protocol}}
                      \alt {\Value \v{}}
                      \alt {\Unionsplice {\overrightarrow{\t{}}}}
                      \alt {\In {\t{}\ \overrightarrow{\t{}}}}
                      \\ &\alt&
                      %\alt (\Not\ \t{})
                      {\All {\overrightarrow{\x{}}} {\t{}}}
                      \alt
                      {\TFn {\overrightarrow{\x{}}} \t{}}
                      \alt
                      {\Arrow {\overrightarrow{ \hastype {\x{}} {\t{}}}}
                                   {\t{}}
                                   {\filterset {\prop{}} {\prop{}}}
                                   {\object{}}}
                      \alt
                      {\ArrayTwo {\t{}} {\t{}}}
                      \alt {\HMapc {\mandatory}}
                      \alt {\HMapp {\mandatory} {\absent}}

@@

\begin{figure*}
\begin{mathpar}
\infer [T-Local]
{ \inpropenv {\propenv{}} {\isprop {\x{}} {\t{}}}}
{ \judgementfillcol {\propenv{}} 
                   {\hastype {\x{}} {\t{}}}
                   {\filterset {{\notprop {\falsydiff} {\x{}}}} {{\isprop {\falsydiff} {\x{}}}}}
                   {\x{}}
                   }

\infer [T-True]
{}
{ \judgementfillcol {\propenv{}}
      {{\hastype {\true{}} {\True{}}}}
      {{\filterset {\topprop{}} {\botprop{}}}}
      {\emptyobject{}}
    }

\infer [T-False]
{}
{ \judgementfillcol {\propenv{}}
      {{\hastype {\false{}} {\False{}}}}
      {\filterset {\botprop{}} {\topprop{}}}
    {\emptyobject{}}
   }

\infer [T-Nil]
{}
{ \judgementfillcol {\propenv{}} 
      {{\hastype {\nil{}} {\Nil{}}}} 
      {\filterset {\botprop{}} {\topprop{}}}
      {\emptyobject{}} 
    }

\infer [T-Do]
{ \judgementfillcol {\propenv{}} {\hastype {\e1} {\t1}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop1}}} {\object{1}} 
\\\\
  \judgementfillcol {\propenv{}, {\trdiff {\orprop {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}}}
                 {\hastype {\e{}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}} 
                 {\object{}}
  }
{ \judgementfillcol {\propenv{}} {\hastype {\doexp {\e1} {\e{}}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}} {\object} 
             }

\infer [T-If]
{ \judgementfillcol {\propenv{}} {\hastype {\e1} {\t{1}}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}
                 {\object{1}} 
             \\\\
  \judgementfillcol {\propenv{}, {\thenprop {\prop{1}}}}
                 {\hastype {\e2} {\t{}}} {\filterset {\thenprop {\prop{2}}} {\elseprop {\prop{2}}}}
                 {\object{}}
  \\\\
  \judgementfillcol {\propenv{}, {\elseprop {\prop{1}}}}
                 {\hastype {\e3} {\t{}}} {\filterset {\thenprop {\prop{3}}} {\elseprop {\prop{3}}}}
                 {\object{}}
             }
{ \judgementfillcol {\propenv{}} {\hastype {\ifexp {\e1} {\e2} {\e3}} {\t{}}} 
                 {\filterset {\orprop {\thenprop {\prop{2}}} {\thenprop {\prop{3}}}} 
                             {\orprop {\elseprop {\prop{2}}} {\elseprop {\prop{3}}}}}
                 {\object{}} }

\infer [T-Let]
{ \judgementfillcol {\propenv{}} {\hastype {\e{1}} {\s{}}} {\filterset {\thenprop {\prop{1}}} {\elseprop {\prop{1}}}}
             {\object{1}}
  \\\\
  \judgementfillcol {\propenv{}, {\isprop {\s{}} {\x{}}}, 
                    {\impprop {\notprop {\falsydiff} {\x{}}} {\thenprop {\prop{1}}}}, 
                    {\impprop {\isprop {\falsydiff} {\x{}}} {\elseprop {\prop{1}}}}}
             {\hastype {\e{}} {\t{}}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
             {\object{}} 
             }
{ \judgementfillcol {\propenv{}} {\hastype {\letexp {\x{}} {\e{1}} {\e{}}} {\t{}}}
             {\replacefor {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                          {\object{1}}
                          {\x{}}}
             {\replacefor {\object{}} 
                          {\object{1}}
                          {\x{}}}
             }

\infer [T-Throw]
{ \judgementfillcol {\propenv{}} {\hastype {\e{}} {\Throwable}} {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
}
{ \judgementfillcol {\propenv{}} {\hastype {\throwexp {\e{}}} {\Bottom}} {\trdiff {\filterset {\botprop{}} {\botprop{}}}}
             {\emptyobject{}} }

\infer [T-App]
{ \judgementfillcol {\propenv{}} {\hastype {\e{}} {\Arrow {\hastype {\x{}} {\s{}}}
                                                       {\t{}}
                                                       {\filterset {\thenprop {\prop{f}}}
                                                                   {\elseprop {\prop{f}}}}
                                                       {\object{f}}}}
                {\filterset {\thenprop {\prop{}}}
                            {\elseprop {\prop{}}}}
                {\object{}}
  \\\\
  \judgementfillcol {\propenv{}}
                 {\hastype {\ep{}} {\s{}}}
                 {\filterset {\thenprop {\propp{}}}
                             {\elseprop {\propp{}}}}
                 {\objectp{}} 
}
{ \judgementfillcol {\propenv{}} {\hastype {\appexp {\e{}} {\ep{}}}
                                        {\replacefor {\t{}}
                                                     {\objectp{}}
                                                     {\x{}}}}
                 {\replacefor {\filterset {\thenprop {\prop{f}}}
                                          {\elseprop {\prop{f}}}}
                              {\objectp{}}
                              {\x{}}}
                 {\replacefor {\object{f}}
                              {\objectp{}}
                              {\x{}}}
}

\infer [T-Abs]
{ \judgementfillcol {\propenv{}, {\isprop {\s{}} {\x{}}}}
            {\hastype {\e{}} {\t{}}}
             {\filterset {\thenprop {\prop{}}}
                         {\elseprop {\prop{}}}}
             {\object{}}}
{ \judgementfillcol {\propenv{}} {\hastype {\e{}} {\ArrowOne {\x{}} {\s{}}
                                                      {\t{}}
                                                      {\filterset {\thenprop {\prop{}}}
                                                                  {\elseprop {\prop{}}}}
                                                      {\object{}}}}
                {\filterset {\topprop}
                            {\botprop}}
                {\object{}}}

\infer [T-Try]
{ \judgementfillcol {\propenv{}} {\hastype {\e{1}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
\\\\
  \judgementfillcol {\propenv{}, {\isprop {\x{}} {\class{1}}}}
             {\hastype {\e{2}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
\\
  \judgementfillcol {\propenv{}}
             {\hastype {\e{3}} {\t{3}}}
                 {\filterset {\thenprop {\prop{3}}} {\elseprop {\prop{3}}}}
                 {\object{3}}
}
{ \judgementfillcol {\propenv{}} {\hastype {\tryexp {\e{1}} {\class{1}} {\x{}} {\e{2}} {\e{3}}} {\t{}}}
                 {\filterset {\thenprop {\prop{}}} {\elseprop {\prop{}}}}
                 {\object{}}
                 }

\infer [T-New]
{ \ctorparams {\overrightarrow {\e{i}}} {\overrightarrow {\t{i}}}
  \\
  {\overrightarrow{
  \judgementfillcol {\propenv{}} {\hastype {\e{i}} {\t{i}}} 
             {\filterset {\prop{}} {\prop{}}}
         {\object{}}}}^{i}
           }
{ \judgementfillcol {\propenv{}} {\hastype {\newexp {\class{}} {\overrightarrow{\e{i}}}} {\class{}}}
             {\filterset {\topprop{}} {\botprop{}}}
             {\emptyobject}
             }

@@

\infer [T-Subsume]
{ \judgementfillcol {\propenv{}} {\hastype {\e{}} {\t{}}}
             {\filterset {\thenprop {\prop{}}}
                         {\elseprop {\prop{}}}}
             {\object}
\\\\
\inpropenv {\propenv{}, {\thenprop {\prop{}}}} {\thenprop {\propp{}}}
\\
\inpropenv {\propenv{}, {\elseprop {\prop{}}}} {\elseprop {\propp{}}}
\\\\
\inpropenv {} {\issubtype {\t{}} {\tp{}}}
\\
\inpropenv {} {\issubtype {\object{}} {\objectp{}}}
}
{ \judgementfillcol {\propenv{}} {\hastype {\e{}} {\tp{}}}
             {\filterset {\thenprop {\propp{}}}
                         {\elseprop {\propp{}}}}
             {\objectp}}
